---
title: "Untitled"
author: "Edwin Thoen"
date: "6/27/2017"
output: html_document
---

## List

Subsetting a list with `[` will always return a list, it can be subsetted on the same five ways as atomic vectors. For pulling out specic objects use `[[` or `$`.

```{r}

```

## Matrices

Most common way is generalised from 1d; use vectors for row and column indexing:

```{r}
x <- matrix(1:9, 3, 3)
colnames(x) <- letters[1:3]

x[c(TRUE, FALSE, TRUE), 1:2]
x[, 3]
x[, "a"]
```

Return of `[` will be of the lowest possible dimensionality. Use `drop = FALSE` to avoid this:

```{r}
x[1:2, 1, drop = FALSE]
```

Because matrices are atomic vectors with a dimensions attribute, we can also do atomic vector subsetting with a single vector. Elements are stored column wise:

```{r}
x[c(3, 2, 8)]
```

Alternatively we can use integer matrices to suset a higher dimension object:

```{r}
x <- matrix(round(rnorm(9), 1), 3, 3)
ind_mat <- matrix(c(1, 1, 3, 2, 3, 2), nrow = 3)
x[ind_mat]
```

## Data Frames

Data frames act like matrices when indexed with two vectors, and like lists when indexed with on vector. Since list subsetting with `[` always returns a list, the following is a one column data frame:

```{r}
x <- head(mtcars, 3)
x["cyl"]
```
However, matrix indexing will always return the lowest possible dimensionality. This is an atomic vector.

```{r}
x[, "cyl"]
```

```{r}
my_diag <- function(x) {
  n <- x %>% dim %>% min
  ind_mat <- matrix(rep(1:n, 2), ncol = 2)
  x[ind_mat]
}
```

## Subsetting Operators

Both `[[` and `$` pull out a single element from a list (also from an atomic vector, but `[` does so too). You can use `[[` with both positive integers and names.

```{r}
a <- list("jos" = 42)
a[["jos"]]
a[[1]]
```

When supplied with a vector, the indexing happens recursively.

## Simplifying vs Preserving

For many objects/types, there is both a simplifying and a preserving way to subset. Preserving always returns the same data type, simplifying depends on the object.

#### Atomic vector
* preserving = `x[1]`
* simplifying = `x[[1]]`, will drop element names

#### List
* preserving = `x[1]`
* simplifying = `x[[1]]`, will return the first element, not a list of length 1

#### Factor
* preserving = `x[1]`
* simplifying = `x[1, drop = TRUE]`, will drop levels not in the selection

#### Matrix or Array
* preserving = `x[1, , drop = FALSE]`
* simplifying = `x[1, ]` abandon all dimension with length 1

#### Data frame
* preserving = `x[, 1, drop = FALSE]`
* simplifying = `x[, 1]` return vector if output is a single column

## $
The `$` is shorthand for `x[[{str}, , exact = FALSE]]`, this means it will do partial matching.

## Missingness / out of bounds
Subsetting with oob indexes and missing values has different behavior for `[` and `[[`, and also different for vectors and lists. See complete list in book, for atomic vectors it works like this:

```{r}
x <- 1:4
x[5]
x[NA]
x[NA_real_]
x[NULL]
```

## Subsetting and assignment

All subsetting can be used for assignment. 

```{r}
x <- 1:5
x[2:3] <- 3:2
```
Left and right should be of equal length:

```{r}
x[2] <- 4:5
```
Although this works due to recycling:

```{r}
x[2:3] <- 4
```
There is no checking for duplicates:

```{r}
x[c(2,2)] <- 4:5
x
```

Subsetting with nothing can be useful when we want to preserve the structure of the original object.

```{r}
a <- mtcars[1:2, 1:2]
a[] <- lapply(a, as.integer)
a
a <- lapply(a, as.integer)
a
```

Assing `NULL` to a list elemnent will drop the element.

```{r}
a["cyl"] <- NULL
a
```
If we want to add a literal `NULL` use `list(NULL)`:
```{r}
a["mpg"] <- list(NULL)
a
```

## Applications

Many things are implemented in easy to use functions. This is how to build it with basic subsetting.

#### Lookup table

Using character matching to find a corresponding element.

```{r}
lookup <- c("the Netherlands", "Germany")
names(lookup) <- c("Nl", "De")
some_char_vec <- sample(c("Nl", "De"), 10, TRUE)
lookup[some_char_vec] %>% unname
```

#### Matching and merging

Instead of a lookup in a vector we want to do a lookup in data frame. We can do this by using `match`, which produces an integer with which we can subset.

```{r}
nl_ge_info <- data.frame(abbr = c("Nl", "De"), 
                         language = c("Dutch", "German"),
                         inhabitants = c(17, 82),
                         type_of_country = c("Awesome", "Boring"))
match(some_char_vec, nl_ge_info$abbr)
nl_ge_info[match(some_char_vec, nl_ge_info$abbr), ] %>% slice(3)
```

Or we can add the lookup names as rownames and do to charachter lookup, like on the names earlier.

```{r}

```


## Exercises

```{r}
permute_both <- function(x) x[sample(1:nrow(x)) ,sample(seq_along(x))]

sample_cont_rows <- function(x, m) {
  stopifnot(is.data.frame(x))
  stopifnot(nrow(x) > m)
  max_start <- nrow(x) - m + 1
  start <- sample(1:max_start, 1)
  x[seq(start, length.out = m), ]
}

x[names(x) %>% sort]

```

