---
title: "Ch 8 Functional Programming"
author: "Edwin Thoen"
date: "8/1/2017"
output: html_document
---

## 8.1 Motivation

When copy-pasting code, we make mistakes easily. Rather we put repetive actions into functions. Duplicating also makes code harder to be maintained or expanded. DRY principle, "Don't Repeat Yourself": every piece of knowledge must have a single, unambiguous, authorative representation within a system. 

`lapply()` is called a **functional**: it takes a function as an argument. Key of FP: write simple functions that can be understood in isolation and can be composed to do something more complex. A **closure** is a function that makes and return function. 

## 8.2 Anonymous functions

In R functions are objects in their own right, there is no special syntax around them. We can work with them without naming them, like in `lapply()`. Just like other functions they have formals, a body and a parent. They are convenient for lambda functions, but making them too complex soon makes them confusing. Start naming them then. Rule of thumb, fit on one line, no {}.

#### 8.2.1

```{r}
match.fun("mean")
```

This doesn't make sense because functions have no special status in R. They can be just assigned to a variable name. The body can even be assigned to multiple names, it does not have to be uniquely attached to a name. 

```{r, eval=FALSE}
lapply(mtcars, function(x) sd(x) / mean(x))
```

## 8.3 Closures

These are functions written by functions. They enclose the environment of their parent and can access all its variables. Gives you two levels of parameters. 

```{r}
pwr <- function(e) function(x) x ^ e
square <- pwr(2)
cube   <- pwr(3)
square
cube
```

Code of the above functions is exactly equal. The difference is in the parent environment, where the `e` parameter is defined. 


